An Empirical Analysis of the Internet Engineering Task Force with Computational Methods

### 2024-09-29 paper.tex
Update §4.1

This paper summarises and expands on our prior work
\cite{mcquistin:2021:characterising,khare:2022:web-we-weave,
healey:2023:power,mcquistin:2023:errare,khare:2023:tracing,
karan:2023:leda,healey:2023:power-frontiers,barnes:2024:temporal}

@@

% The following is from our IMC 2021 paper, Section 3.2:

The IETF Datatracker maintains metadata about document authors, including
their names, email addresses, affiliations, and location information. This
dataset does not cover the entire RFC corpus, with metadata available for
authors of RFCs published from 2001, and where it has been provided.
Country data is available for around 70\% of authors, while affiliation
information is given for around 80\%. In this section, we look at the
authorship of all of the RFCs published each year, using the available
metadata. An author is counted once in a year for each affiliation or
location they hold.

%--------------------------------------------------------------------------------------------------
\subsection{Geographic Distribution of Authors}
Figures~\ref{fig:author_countries_normalised} and
\ref{fig:author_continents_normalised} show the proportion of authors from
countries and continents, respectively. The IETF has signalled that it
wishes to encourage greater geographical diversity~\cite{rfc7704,
ietfblog:diversity}. Without an explicit goal, we frame our findings within
the context of global population distribution.  We find that North America,
while still disproportionately over represented, is becoming less dominant.
75\% of authors were from North America in 2001, and this has declined to
44\% in 2020. At the same time, representation of both Europe and Asia has
grown, from 17\% to 40\% and 6\% to 14\%, respectively.  However, Africa
and South America remain heavily underrepresented, with only $\approx$0.5\%
of authors coming from either continent in 2020.  This suggests that, if
the IETF is to become more geographically representative, further efforts
are needed.

@@

%--------------------------------------------------------------------------------------------------
\subsection{Author Affiliations}
Figure~\ref{fig:author_affiliations_normalised} shows the top 10
affiliations by proportion of authors each year. Affiliation data is
gathered from the Datatracker, and processed to normalise affiliation
names, removing common variations in spelling, and amalgamating known
subsidiaries and merged companies. For example, Huawei and Futurewei are
combined as Huawei, and Sun Microsystems is merged with Oracle.

>>>>

This paper summarises and expands on our prior work
\cite{mcquistin:2021:characterising,khare:2022:web-we-weave,
healey:2023:power,mcquistin:2023:errare,khare:2023:tracing,
karan:2023:leda,healey:2023:power-frontiers,barnes:2024:temporal}.

@@

In this section, we look at the authorship of published RFCs to explore
how participation in the IETF has changed over time. We show that the
geographic distribution of RFC authors has shifted over time, with the
proportion of authors from the US decreasing and a corresponding rise in
authors from Europe and China (\S\ref{sec:trends-demographics:geographic}).
We similarly explore shifts in affiliation, showing increasing influence of
Huawei and ongoing shifts in the fortunes of many prominent US and European
technology companies (\S\ref{sec:trends-demographics:affiliation}). The
ongoing influence of Cisco and strong input from academic institutions is
also noted.

% The following is from our IMC 2021 paper, Section 3.2.

%--------------------------------------------------------------------------------------------------
\subsection{Geographic Distribution of Authors}
\label{sec:trends-demographics:geographic}

The IETF Datatracker maintains metadata about document authors, including
their names, email addresses, affiliations, and location information. This
dataset does not cover the entire RFC corpus, with metadata available for
authors of RFCs published from 2001, and where it has been provided country
data is available for around 70\% of authors, while affiliation information
is provided for around 80\% of authors.

Figure~\ref{fig:author_countries_normalised} shows how the proportion of
RFC authors from different countries varies over time, with Figure
\ref{fig:author_continents_normalised} breaking this down by continent.
The IETF has signalled that it wishes to encourage greater geographical
diversity~\cite{RFC7704,ietfblog:diversity}.  Without an explicit goal, we
frame our findings within the context of global population distribution.
We find that North America, while still disproportionately over represented,
is becoming less dominant.  75\% of authors were from North America in 2001,
and this has declined to 44\% in 2020. At the same time, representation of
both Europe and Asia has grown, from 17\% to 40\% and 6\% to 14\%, respectively.
However, Africa and South America remain heavily under-represented, with
only $\approx$0.5\% of authors coming from either continent in 2020.  This
suggests that, if the IETF is to become more geographically representative,
further efforts are needed.

@@

%--------------------------------------------------------------------------------------------------
\subsection{Author Affiliations}
\label{sec:trends-demographics:affiliation}

Figure~\ref{fig:author_affiliations_normalised} shows the top 10
affiliations by proportion of authors each year. Affiliation data is
gathered from the Datatracker, and processed to normalise affiliation
names, removing common variations in spelling, and amalgamating known
subsidiaries and merged companies. For example, Huawei and Futurewei are
combined as Huawei, and Sun Microsystems is merged with Oracle.
